---
title: "Phase 2 Final"
author: "Owen Gillespie"
date: "2025-06-28"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(naniar)
library(readr)
library(ranger)
library(caret)
library(mice)
library(VIM)
library(glmnet)
library(GGally) 
library(ggcorrplot) 
library(MASS) 
library(leaps) 
library(lmtest) 
library(splines)
library(tidymodels)
library(dplyr)
library(rlang)
library(rpart) 
library(rpart.plot) 
library(RColorBrewer) 
library(rattle)
library(e1071)
library(ROCR)
```
## Loading Data
```{r}
train <- read_csv("C:/Users/oweng/Desktop/UNCW/MIS 502/Final Project/train.csv")
summary(train)
test <- read_csv("C:/Users/oweng/Desktop/UNCW/MIS 502/Final Project/test.csv")
train <- train %>%
  mutate_if(is_character,as_factor)
test <- test %>%
  mutate_if(is_character,as_factor)
```

## Dropping Loading Missing Data
```{r}
train <- train%>%
  drop_na(loading)
```

## imputing other missing data
```{r}
imputed_data = mice(train, method = "pmm", m = 5)
train = complete(imputed_data)
gg_miss_var(train) #Missing Data Validation
```

# Splitting the Data
```{r}
set.seed(123)
data_split = initial_split(train, prop = 0.8, strata = failure)
train_set = training(data_split)
test_set = testing(data_split)
```
## Folds
```{r}
set.seed(123)
rf_folds = vfold_cv(train, v = 5)
```


##Regression 
```{r}
train_recipe = recipe(failure ~ ., data = train)
train_model <- linear_reg() %>%
  set_engine("lm")
lm_wflow = 
  workflow() %>% 
  add_model(train_model) %>% 
  add_recipe(train_recipe)

```

## Random Forrest
```{r}
train_recipe = recipe(failure ~ ., data = train)%>%
   step_dummy(all_nominal(), -all_outcomes())
forrest_model <- rand_forest() %>%
  set_engine("ranger")%>%
  set_mode("classification")
forrest_wflow = 
  workflow() %>% 
  add_model(forrest_model) %>% 
  add_recipe(train_recipe)

train_fit = fit(forrest_wflow, train)

trainpredrf = predict(train_fit, train)
head(trainpredrf)

confusionMatrix(trainpredrf$.pred_class, train$failure, positive = "Yes")
## Accuracy @ 0.9936

```
##Testing the Random Forrest
```{r}
testpredrf = predict(train_fit, test_set)
head(testpredrf)
confusionMatrix(testpredrf$.pred_class, test_set$failure, 
                positive = "Yes")
##Accuracy .9926
```


## Classification Tree
```{r}
train_recipe

class_tree_model = decision_tree() %>% 
  set_engine("rpart", model = TRUE) %>%
  set_mode("classification")

train2_wflow = 
  workflow() %>% 
  add_model(class_tree_model) %>% 
  add_recipe(train_recipe)

train2_fit = fit(train2_wflow, train)

train2_fit %>%
  extract_fit_parsnip() %>% #error code requested use of extract fit parsnip in place of pull_wkflow
  pluck("fit") 

tree = train2_fit %>% 
  extract_fit_parsnip() %>% 
  pluck("fit")
rpart.plot(tree)

train2_fit$fit$fit$fit$cptable
```


#Testing the Classification Tree
## Only one variable, 21% of the products failed. No split in this tree. 
```{r}
failure_fit_test = fit(train2_wflow, data = test_set)
tree_test = failure_fit_test %>%
  extract_fit_parsnip() %>%
  pluck("fit")
rpart.plot(tree_test)
```

#Logistic Regression
```{r}
logreg_model = 
  logistic_reg() %>% 
  set_engine("glm") 

train_recipe

logreg_wf = workflow() %>%
  add_recipe(train_recipe) %>% 
  add_model(logreg_model)

train3_fit = fit(logreg_wf, train)

summary(train3_fit$fit$fit$fit)

logreg_fit = fit(logreg_wf, data = train)

logregpred = predict(train3_fit, train, type="prob")

head(logreg_fit)

```
#Lasso
```{r}
lasso_model =
  logistic_reg(mixture = 1,penalty = 0.01) %>% 
  set_engine("glmnet") 

train_lasso_recipe = recipe(failure ~ ., data = train) %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_center(all_predictors()) %>% 
  step_scale(all_predictors()) 

lasso_wflow =
  workflow() %>% 
  add_model(lasso_model) %>% 
  add_recipe(train_lasso_recipe)
lasso_fit = fit(lasso_wflow, train)

lasso_fit %>%
  pull_workflow_fit() %>%
  pluck("fit")  
```

#Test Set Data Missing
```{r}
gg_miss_var(test)
```
```{r}
imputed_test = mice(test, method = "pmm", m = 5)
library(dplyr) 
test_complete = complete(imputed_test)
gg_miss_var(test_complete)

testpredrf <- predict(train_fit, new_data = test)

test_predictions = as_tibble(testpredrf)
test_predictions = test %>% 
  dplyr::select(id) %>%
  bind_cols(testpredrf) 

colnames(test_predictions) <- c("id", "failure")

test_with_predictions <- bind_cols(test, test_predictions)
write_csv(test_predictions, "submission.csv")
```

## Visualization 1
```{r}
test_with_predictions %>%
  count(product_code, failure) %>%
  ggplot(aes(x = product_code, y = n, fill = failure)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("skyblue","red"))+
  geom_text(aes(label = n),position = position_dodge(width = 0.8),vjust = -0.5, size = 3) +
  labs(title = "Predicted Failure Count by Product Code",x = "Product Code",y = "Count",fill = "Failure Status") +
  theme_classic()
```
## Visualization 2

```{r}
ggplot(test_with_predictions, aes(x = product_code, y = loading, fill = failure)) +
  geom_boxplot(position = position_dodge(0.8)) +
  labs(title = "Loading by Product Code and Failure", y = "Loading", x = "Product Code") +
  scale_fill_manual(values = c("skyblue","red"))+
  theme_classic()
```
## Visualization 3
```{r}
ggplot(test_with_predictions, aes(x = failure, fill = failure)) +
  geom_bar() +
  labs(title = "Predicted Failed vs Passed Products", x = "Failure Status", y = "Count", fill = "Failure") +
  geom_text(stat = "count", 
            aes(label = paste0(round(..count../sum(..count..) * 100, 0), "%")),
            vjust = -.5)+
  scale_fill_manual(values = c("No" = "skyblue", "Yes" = "red")) +
  theme_classic()

```

## Visulazation 4
```{r}
ggplot(test_with_predictions, aes(x = loading, fill = failure)) +
  geom_histogram(binwidth = 3, position = "identity", alpha = 0.3) +
  scale_fill_manual(values = c("skyblue", "red")) +
  labs(title = "Predicted Distribution of Loading Values by Failure Status",
       x = "Loading", y = "Count", fill = "Failure") +
  theme_classic()
```
## Visualization 5
```{r}
ggplot(test_with_predictions, aes(x = loading, fill = failure)) +
  geom_histogram(position = "identity", alpha = 0.3, bins = 30) +
  facet_wrap(~ product_code) +
  scale_fill_manual(values = c("skyblue", "red")) +
  labs(title = "Predicted Loading Distribution by Product Code and Failure Status",
       x = "Loading", y = "Count", fill = "Failure") +
  theme_classic()
```
```{r}
test_with_predictions %>%
  ggplot(aes(x = product_code, y = loading, color = failure)) +
  geom_jitter(width = 0.3, alpha = 0.3) +
  scale_color_manual(values = c("No" = "skyblue", "Yes" = "red"))+
  labs(title = "Scatterplot of Predicted Loading by Product and Failure", x = "Product Code", y = "Loading", fill = "Failure") +
  theme_classic()
```
```{r}
ggplot(test_with_predictions, aes(x = measurement_2, fill = failure)) +
  geom_histogram(position = "identity", alpha = 0.3, bins = 25) +
  scale_fill_manual(values = c("skyblue", "red")) +
  labs(title = "Measurement 2 Distribution by Product Code and Failure Status",
       x = "Measurement_2", y = "Count", fill = "Failure") +
  theme_classic()
```
```{r}
ggplot(test_with_predictions, aes(x = measurement_15, fill = failure)) +
  geom_histogram(position = "identity", alpha = 0.3, bins = 25) +
  scale_fill_manual(values = c("skyblue", "red")) +
  labs(title = "Measurement 15 Distribution by Product Code and Failure Status",
       x = "Measurement_15", y = "Count", fill = "Failure") +
  theme_classic()
```
```{r}
ggplot(test_with_predictions, aes(x = measurement_17, fill = failure)) +
  geom_histogram(position = "identity", alpha = 0.3, bins = 25) +
  scale_fill_manual(values = c("skyblue", "red")) +
  labs(title = "Measurement 17 Distribution by Product Code and Failure Status",
       x = "Measurement_17", y = "Count", fill = "Failure") +
  theme_classic()
```
```{r}
ggplot(test_with_predictions, aes(x = measurement_2, fill = failure)) +
  geom_bar(position = "dodge") +
  geom_text(stat = "count",
    aes(label = paste0(round(after_stat(count / sum(count) * 100), 0), "%")),
    position = position_dodge(width = 0.9),vjust = -0.5,size = 3) +
  labs(title = "Failed vs Passed Products by Measurement 2", x = "Measurement 2", y = "Count", fill = "Failure") +
  scale_fill_manual(values = c("No" = "skyblue", "Yes" = "red")) +
  theme_classic()
```


```

```

