---
title: "Phase 1"
author: "Owen Gillespie"
date: "2025-06-15"
output:
  word_document: default
  pdf_document: default
---
## Library Download / Data Download
```{r}
library(tidyverse)
library(naniar)
library(readr)
getwd()
setwd("C:/Users/oweng/Desktop/UNCW/MIS 502/Final Project")
train <- read_csv("C:/Users/oweng/Desktop/UNCW/MIS 502/Final Project/train.csv")
```

## Missing Data
```{r}
gg_miss_var(train)
train_loading_na <- train %>%
  drop_na(loading)
```
```{r}
library(dplyr)
train <- train %>%
mutate(across(where(is_character), as_factor))
```


## Visualization 1 - Product Codes that Failed vs Pass
```{r}
train_counts <- train %>%
  count(product_code, failure)

train_percent <- train %>%
    count(product_code, failure) %>%
  group_by(product_code) %>%
  mutate(percent = n / sum(n) * 100)

train %>%
  count(product_code, failure) %>%
  ggplot(aes(x = product_code, y = n, fill = failure)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("skyblue","red"))+
  geom_text(aes(label = n),position = position_dodge(width = 0.8),vjust = -0.5, size = 3) +
  labs(title = "Failure Count by Product Code",x = "Product Code",y = "Count",fill = "Failure Status") +
  theme_classic()

```
# Visualization 2 - Loading and Failure
```{r}
ggplot(train_loading_na, aes(x = product_code, y = loading, fill = failure)) +
  geom_boxplot(position = position_dodge(0.8)) +
  labs(title = "Loading by Product Code and Failure", y = "Loading", x = "Product Code") +
  scale_fill_manual(values = c("skyblue","red"))+
  theme_classic()
```
## Visulaization 3 - total pass / fail

```{r}
failure_counts <- train %>%
  group_by(failure) %>%
  summarise(count = n()) %>%
  mutate(percent = round(count / sum(count)*100, 0))

ggplot(train, aes(x = failure, fill = failure)) +
  geom_bar() +
  labs(title = "Failed vs Passed Products", x = "Failure Status", y = "Count", fill = "Failure") +
  geom_text(stat = "count", 
            aes(label = paste0(round(..count../sum(..count..) * 100, 0), "%")),
            vjust = -.5)+
  scale_fill_manual(values = c("No" = "skyblue", "Yes" = "red")) +
  theme_classic()
failure_counts$count
```

## Visualization 4 - Histogram comparing distribution of loading and product code.

```{r}
ggplot(train, aes(x = loading, fill = failure)) +
  geom_histogram(binwidth = 3, position = "identity", alpha = 0.3) +
  scale_fill_manual(values = c("skyblue", "red")) +
  labs(title = "Distribution of Loading Values by Failure Status",
       x = "Loading", y = "Count", fill = "Failure") +
  theme_classic()
```
##Visualization 5 histograms per product code based on loading that passed and failed
```{r}
ggplot(train, aes(x = loading, fill = failure)) +
  geom_histogram(position = "identity", alpha = 0.3, bins = 30) +
  facet_wrap(~ product_code) +
  scale_fill_manual(values = c("skyblue", "red")) +
  labs(title = "Loading Distribution by Product Code and Failure Status",
       x = "Loading", y = "Count", fill = "Failure") +
  theme_classic()
```
##Visualization 6 - Scatter plot of failure on loading per product code
```{r}
train %>%
  ggplot(aes(x = product_code, y = loading, color = failure)) +
  geom_jitter(width = 0.3, alpha = 0.3) +
  scale_color_manual(values = c("No" = "skyblue", "Yes" = "red"))+
  labs(title = "Scatterplot of Loading by Product and Failure", x = "Product Code", y = "Loading", fill = "Failure") +
  theme_classic()
```
##Visualization 7 - Density plot per product code 
```{r}
train %>%
  ggplot(aes(x = loading, fill = failure)) +
  geom_density(alpha = 0.3) +
  facet_wrap(~product_code)+
  scale_fill_manual(values = c("skyblue", "red"))+
  labs(title = "Density Plot of Loading by Failure Status", x = "Loading", y = "Density", fill = "Failure") +
  theme_minimal()
```
##Visualization 8 - Attribute 2 Grouped Bar Pass / Fail
```{r}
ggplot(train, aes(x = attribute_2, fill = failure)) +
  geom_bar(position = "dodge") +
  geom_text(stat = "count",
    aes(label = paste0(round(after_stat(count / sum(count) * 100), 0), "%")),
    position = position_dodge(width = 0.9), vjust = -0.5,size = 3) +
  labs(title = "Failed vs Passed Products by Attribute 2", x = "Attribute 2", y = "Count", fill = "Failure") +
  scale_fill_manual(values = c("No" = "skyblue", "Yes" = "red")) +
  theme_classic()
```
##Visualization 9 - Attribute 3 Grouped Bar Pass / Fail
```{r}
ggplot(train, aes(x = attribute_3, fill = failure)) +
  geom_bar(position = "dodge") +
  geom_text(stat = "count",
    aes(label = paste0(round(after_stat(count / sum(count) * 100), 0), "%")),
    position = position_dodge(width = 0.9),vjust = -0.5,size = 3) +
  labs(title = "Failed vs Passed Products by Attribute 3", x = "Attribute 3", y = "Count", fill = "Failure") +
  scale_fill_manual(values = c("No" = "skyblue", "Yes" = "red")) +
  theme_classic()
```





```




